---
title: "Plotly and Flexdashboard"
output: 
  html_document:
    toc: true
    toc_float: true
---

Plotly is a flexible framework for producing interactive graphics; it has a variety of implementations, including one for R. We’ll take a look at a few common plot types, and then introduce `flexdashboards` as a way to collect plots (either static or interactive).

```{r}
library(tidyverse)
library(p8105.datasets)

library(plotly)
```

We’re going to focus on the Airbnb data. The code below extracts what we need right now; specifically, we select only a few of the variables and filter to include a subset of the data. In part, this makes sure that the resulting dataset and plots are computationally feasible – for large datasets, you may need to downsample.
```{r}
data(nyc_airbnb)

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(rating = review_scores_location / 2) |>
  select(
    neighbourhood_group, neighbourhood, rating, price, room_type, lat, long) |>
  filter(
    !is.na(rating), 
    neighbourhood_group == "Manhattan",
    room_type == "Entire home/apt",
    price %in% 100:500)
```

# Plotly Scatterplot
There are several practical differences comparing `ggplot` and `plot_ly`, but the underlying conceptual framework is similar. We need to define a dataset, specify how variables map to plot elements, and pick a plot type.

Below we’re plotting the location (latitude and longitude) of the rentals in our dataset, and mapping `price` to color. We also define a new variable `text_label` and map that to text.

The type of plot is `scatter`, which has several “modes”: `markers` produces the same kind of plot as `ggplot::geom_point`, `lines` produces the same kind of plot as `ggplot::geom_line`.

```{r}
nyc_airbnb |>
  mutate(text_label = str_c("Price: $", price, "\nRating: ", rating)) |> 
  plot_ly(
    x = ~lat, y = ~long, type = "scatter", mode = "markers",
    color = ~price, text = ~text_label, alpha = 0.5)
```

This can be a useful way to show the data – it gives additional information on hovering and allows you to zoom in or out, for example.

# Plotly Boxplot
Next up is the boxplot. The process for creating the boxplot is similar to above: define the dataset, specify the mappings, pick a plot type. Here the type is `box`, and there aren’t modes to choose from.

```{r}
nyc_airbnb |> 
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) |> 
  plot_ly(y = ~price, color = ~neighbourhood, type = "box", colors = "viridis")
```

Again, this can be helpful – we have a five-number summary when we hover, and by clicking we can select groups we want to include or exclude.

# Plotly Barchart
Lastly, we’ll make a bar chart. Plotly expects data in a specific format for bar charts, so we use `count` to get the number of rentals in each neighborhood (i.e. to get the bar height). Otherwise, the process is pretty familiar.

```{r}
nyc_airbnb |> 
  count(neighbourhood) |> 
  mutate(neighbourhood = fct_reorder(neighbourhood, n)) |> 
  plot_ly(x = ~neighbourhood, y = ~n, color = ~neighbourhood, type = "bar", colors = "viridis")
```

# `ggplotly`

You can convert a `ggplot` object straight to an interactive graphic using `ggplotly`. For example, the code below recreates our scatterplot using `ggplot` followed by `ggplotly`.

```{r}
scatter_ggplot = 
  nyc_airbnb |>
  ggplot(aes(x = lat, y = long, color = price)) +
  geom_point(alpha = 0.25) +
  coord_cartesian()

ggplotly(scatter_ggplot)
```

We can recreate our boxplot in a similar way.

```{r}
box_ggplot = 
  nyc_airbnb |> 
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) |> 
  ggplot(aes(x = neighbourhood, y = price, fill = neighbourhood)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplotly(box_ggplot)
```

If I really want an interactive plot to look good, I’ll use `plot_ly` to build it – `ggplot` was designed with static plots in mind, and the formatting and behavior of `ggplotly` is less visually appealing (to me) than `plot_ly`.

I use `ggplot` for static plots, and I make static plots way, way more frequently than interactive plots. Sometimes I’ll use `ggplotly` on top of that for some quick interactivity; this can be handy to do some zooming or inspect outlying features.

# `flexdashboard`
Clearly you can embed interactive graphics in HTML files produced by R Markdown; this is a handy time to introduce dashboards. In short, dashboards are a collection of related graphics (or tables, or other outputs) that are displayed in a structured way that’s easy to navigate.

You can create dashboards using the `flexdashboard` package by specifying `flex_dashboard` as the output format in your R Markdown YAML. There are a variety of layout options, but we’ll focus on a pretty simple structure produced by the template below. This is the default dashboard template in R Studio – if you have `flexdashboard` installed, you can use `File > New File > R Markdown > From Template` to create a new .Rmd file with the standard structure below. This dashboard has space for three plots. 

Dashboard layouts are controlled by specifying columns and rows, and potentially subdividing these. We specified a two-column layout with set column widths, and then divided the second column into two panels. Using tabbed browsing and multiple pages can also be really useful.

# `flexdashboard`'s on Websites

You can share the HTML files for dashboards directly (e.g. by email); you can also host these online to make the dashboard visible to others. That process is essentially the same as for any other website you’d make.

However, the website’s `_site.yml` file conflicts with the dashboard’s `YAML` header regarding the output format – and the website’s `_site.yml` “wins”. To address this issue, instead of knitting you can use this command to knit the dashboard.

```{r, echo = FALSE, eval = FALSE}
rmarkdown::render("dashboard_template.Rmd", output_format = "flexdashboard::flex_dashboard")
```

This will create `dashboard_template.html` but not open it in RStudio’s Viewer pane; you can open the file in a browser instead. Alternatively, using RStudio’s Build pane to Build Website will produce the same results. To illustrate, we’ll put the dashboard we just created on a website for this topic.

All of this `YAML` business is only an issue for dashboards embedded in websites; a standalone dashboard (in a non-website GH repo or R Project) can be knit using the same process as other .Rmd files.








