---
title: "Data Science in Action"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(plotly)

library(tidyverse)
library(p8105.datasets)
```

```{r}
weather = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX", "SNOW", "SNWD"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |> 
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10, 
    snwd = round((snwd/25.4), digits = 2)
  )

```

!!!!!! Need to separate out date variable and add a month variable !!!


Column {data-width=650}
-----------------------------------------------------------------------

### Snow Depth in Waterhole vs. Central Park, 2021 - 2022

!!! Would be cool to have this color filled in - google how to do that!!!

```{r}
weather |> 
  filter(name != "Molokai_HI") |> 
  mutate(
    text_label = str_c("Snow Depth (in): ", snwd, "\nDate: ", date)
  ) |> 
  plot_ly(
    x = ~date, y = ~snwd, type = "scatter", mode = "line", 
    color = ~name, colors = "viridis", text = ~text_label, alpha = 0.5
  ) |>
  layout(
    xaxis = list(title = "Date"), 
    yaxis = list(title = "Snow Depth (in)"), 
    legend  = list(title = list(text = "Location"))
  )
```

Column {data-width=350}
-----------------------------------------------------------------------

### Total Precipitation by Month, 2021 - 2022

Fix months so that they're in order of date rather than alphabetical

```{r}
weather |> 
  separate(date, into = c("year", "month", "day"), sep = "-") |> 
  mutate(
    month = as.factor(case_match(month,
                                 "01" ~ "Jan",
                                 "02" ~ "Feb",
                                 "03" ~ "Mar",
                                 "04" ~ "Apr",
                                 "05" ~ "May",
                                 "06" ~ "Jun",
                                 "07" ~ "Jul",
                                 "08" ~ "Aug",
                                 "09" ~ "Sep",
                                 "10" ~ "Oct",
                                 "11" ~ "Nov",
                                 "12" ~ "Dec")),
    month = factor(month, levels = )
  )
  
  
  
  
  group_by(name, month) |> 
  summarize(total_precip = sum(prcp)) |> 
  plot_ly(
    x = ~month, y = ~total_precip, color = ~name, type = "bar", colors = "viridis")
  

```

  mutate(
    response = factor(response, 
                      levels = c("Poor", "Fair", "Good", "Very Good", "Excellent"),

  mutate(
    month = case_match(month,
      "01" ~ "Jan",
      "02" ~ "Feb",
      "03" ~ "Mar",
      "04" ~ "Apr",
      "05" ~ "May",
      "06" ~ "Jun",
      "07" ~ "Jul",
      "08" ~ "Aug",
      "09" ~ "Sep",
      "10" ~ "Oct",
      "11" ~ "Nov",
      "12" ~ "Dec")
  ) |> 


### Chart C

Scatterplot of min (x) and max (y) temp correlation?

```{r}

```

